Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-23 paper.tex
edits in section 3

  Value typing \fbox{$v : T$}
  \begin{gather*}
  \dots \qquad
  \inference{
    v : I
  }{
    \vOOcast{v}{\cOOcast{I}{l}{\TOOdyn}} : \TOOdyn
  }
  \quad
  \inference{
    v : P_1 &
    P_1 \smile P_2
  }{
    \vOOcast{v}{\cOOcast{P_1}{l}{P_2}} : P_2
  }
  \end{gather*}
        \caption{Definition of machine state and auxilliary data
          structures.}
        \label{fig:state}
\end{figure}

@@

  \fbox{$applyCast(v,c) = r$}
  \[
  \begin{array}{rclr}
  \funrule{
    applyCast(v,\cOOcast{\star}{l}{\star})
  }{
    \rOOsucc{v}
  }{}
  \funrule{
    applyCast(\vOOcast{v}{\cOOcast{P_1}{l_1}{\star}},\cOOcast{\star}{l_2}{P_2})
  }{
    applyCast(v,\cOOcast{P_1}{l_2}{P_2})
  }{}
  \funrule{
    applyCast(v,\cOOcast{P}{l}{\star})
  }{
    \rOOsucc{\vOOcast{v}{\cOOcast{P}{l}{\star}}}
  }{}
  \funrule{
    applyCast(v,\cOOcast{P_1}{l}{P_2})
  }{
    \rOOsucc{\vOOcast{v}{\cOOcast{P_1}{l}{P_2}}}
  }{\sidecond{P_1 \smile P_2}}
  \funrule{
    applyCast(v,\cOOcast{P_1}{l}{P_2})
  }{
    \rOOfail{l}
  }{\sidecond{\neg P_1 \smile P_2}}

@@

  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{?^l}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1}{\bot^l}
  }{
    \sidecond{
      \neg \; m_1 \smile m_2
      \; \text{and} \;
      \forall l. t \neq \bot^{l}
    }
  }

\comprulel{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{\epsilon}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1}{\bot^l}
}{
\sidecond{
  \neg \; m_1 \smile m_2
  \; \text{and} \;
  \forall l. t \neq \bot^{l}
}
}
\comprule{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{?^l}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1 \fatsemi^{l} m_2}{t_2}
}{
\sidecond{
  m_1 \smile m_2
  \; \text{and} \;
  \forall l. t \neq \bot^{l}
}
}
  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{\epsilon}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1 \fatsemi^{\ell} m_2}{t_2}
  }{
    \sidecond{
      m_1 \smile m_2
      \; \text{and} \;
      \forall l. t \neq \bot^{l}
    }
  }
%
% \comprule{
%   \hyperCoercionC{h}{m_1}{t_1}
% }{
%   \hyperCoercionC{\epsilon}{m_2}{t_2}
% }{
%   \hyperCoercionC{h}{m'}{t'}
% }{
%   \sidecond{
%     m_1 \fatsemi^{\ell} (m_2, t_2) = (m', t')
%      \; \text{and} \;
%      \forall l. t \neq \bot^{l}
%   }
% }
% 
% \comprule{
%   \hyperCoercionC{h}{m_1}{t_1}
% }{
%   \hyperCoercionC{?^{l'}}{m_2}{t_2}
% }{
%   \hyperCoercionC{h}{m'}{t'}
% }{
%   \sidecond{
%     m_1 \fatsemi^{l'} (m_2, t_2) = (m', t')
%     \; \text{and} \;
%     \forall l. t \neq \bot^{l} 
%   }
% }
  \end{array}
  \]

@@

\figref{fig:HC-D} defines functions that construct \lazyD{} hypercoercions. We 
expect language implementations to call $id(T)$, which constructs an 
identity, $seq(c,c)$, which composes two hypercoercions, and $
cast(T_1,l,T_2)$, which translate a cast to a hypercoercion. Other functions 
in the figure are their helpers.

Let $\ell$ range over $\epsilon$ and labels.
$ c_1 \fatsemi^\ell c_2$ composes hypercoercion $c_1$ and hypercoercion $c_2$.
When both $c_1$ and $c_2$ are $\hyperCoercionI$, their 
composition is also $\hyperCoercionI$.
When $c_1$ is $\hyperCoercionI$ but $c_2$ is not, the head of the 
composition must be a projection. In this 
case, if the head of $c_2$ is a projection, we reuse it. Otherwise, we need a 
label to build the projection. Since the head of $c_2$ is $\epsilon$, its 
source type must be a pretype. Thus we know $\ell$ must be a label and put it 
in the projection. 
When $c_1$ ends with a failure, the composition is $c_1$ itself. In all 
remaining cases, we shall assume that $c_1$ does not end with 
a failure.
When $m_1$ and $m_2$ have different top constructors ($\neg\;m_1 \smile 
m_2$), proceeding to $m_2$ can not make sense. So we end the composition 
with a failure. To do so we need a label. When $c_2$ starts with a 
projection, we shall accuse it of casting a value to a shallowly inconsistent 
type. When $c_2$ starts with a no-op, $\ell$ must be a label. And we accuse 
it of composing shallowly inconsistent hypercoercions. The last two cases 
compose $m_1$ and $m_2$ with a helper function $m \fatsemi^\ell m$, which 
assumes its inputs have the same top constructor. The definition of $m 
\fatsemi^\ell m$ is straightforward. Going back to the last two cases of $c_1 
\fatsemi^\ell c_2$. When $c_2$ starts with a projection, we have no clue 
whether the target type of $m_1$ is the same as the source of $m_2$, 
although we know they are shallowly consistent. So we give the $l$ in the 
projection to $m \fatsemi^\ell m$. When $c_2$ starts with a no-op, we reuse 
the $\ell$.

$seq(c_1,c_2)$ requires that the target type of $c_1$ is the same as the 
source of $c_2$. $c \fatsemi^\ell c$ does the job for it.
$id(T)$ constructs an identity coercion of $T$ with the help of $id(P)$. 
Their definitions are straightforward.
$cast(T_1,l,T_2)$ constructs a coercion from a source type, a label, and a 
target type. $c \fatsemi^\ell c$ does the most job.

\begin{proposition}[\lazyD\ hypercoercion is a monoid]
  \label{thm:hc-monoid}
  For all $c : T_1 \Longrightarrow T_2$,
  $c_1 : T_1 \Longrightarrow T_2$,
  $c_2 : T_2 \Longrightarrow T_3$, and
  $c_3 : T_3 \Longrightarrow T_4$,
  \begin{enumerate}
    \item $seq(id(T_1),c) = c$,
    \item $seq(c,id(T_2)) = c$, and
    \item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
  \end{enumerate}
\end{proposition}
\begin{proof} Part (1) and (2) are by induction on $c$. Part (3) is by 
induction on $c_2$ and case analysis on $c_1$ and $c_3$.
\end{proof}

\begin{proposition}[]
  For all $T$ and $l$, $cast(T,l,T) = id(T) $
\end{proposition}
\begin{proof}
  By induction on T.
\end{proof}

@@

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)$] constructs an identity cast from a type
  \item[$seq(c_1,c_2)$] composes two casts
  \item[$cast(T_1,l,T_2)$] constructs a cast from $T_1$ to $T_2$
  \item[$ applyCast(v,c)$] applies a cast to a value, producing a result
  \end{description}
  We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

@@

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item if $v : T$, then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item if $\judgeType{v}{T_1}$,
    $ \judgeTypeFT{c_1}{T_1}{T_2}$, and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} >>= f & = & f(v) \\
      \rOOfail{l} >>= f & = & \rOOfail{l}
    \end{array}
    \]
  \item if $v : T_1$ and $\neg T_1 \smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item if $v : \star$, 
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item if $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item if $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item if $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
  \item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

\funrule{
  appS(
    \hcvOOcons{v_1}{c_1}{v_2}{c_2},
    \lsOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}
}{
  \rOOsucc{(\hcvOOcons{v_1}{c_1'}{v_2}{c_2'})}
}{
\\ &&
\text{where} \;
c_1' = seq(c_1,cast(T_1,l,T_3))
\\ &&
\text{and} \;
c_2' = seq(c_2,cast(T_2,l,T_4))
}
\funrule{
  appS(\hcvOOinl{v}{c},\lsOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}})
}{
  \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})}
}{}
\funrule{
  appS(\hcvOOinr{v}{c},\lsOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}})
}{
  \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})}
}{}
\funrule{
  appS(v,\cOOcast{P_1}{l}{P_2})
}{
  \rOOfail{l}
}{\sidecond{\neg P_1 \smile P_2}}

@@

  \fbox{$applyMiddle(v,\ell,m) = r$}
  \[
  \begin{array}{rclr}
  \funrule{applyMiddle(\hcvOOtt,\ell,\POOunit)}{\rOOsucc{\hcvOOtt}}{}
  \funrule{applyMiddle(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}\ell,\POOfun{c_3}{c_4})}{
    \rOOsucc{(\hcvOOfun{(c_3 \fatsemi^\ell c_1)}{\rho}{x}{e}{(c_2 
    \fatsemi^\ell c_4)})}
  }{}
  \funrule{applyMiddle(\hcvOOcons{v_1}{c_1}{v_2}{c_2},\ell,\POOprod{c_3}{c_4})}{
    \rOOsucc{(\hcvOOcons{v_1}{(c_1 \fatsemi^\ell c_3)}{v_2}{(c_2 
    \fatsemi^\ell c_4)})}
  }{}
  \funrule{applyMiddle(\hcvOOinl{v}{c_1},\ell,\POOsum{c_3}{c_4})}{
    \rOOsucc{(\hcvOOinl{v}{(c_1 \fatsemi^\ell c_3)})}
  }{}
  \funrule{applyMiddle(\hcvOOinr{v}{c_2},\ell,\POOsum{c_3}{c_4})}{
    \rOOsucc{(\hcvOOinr{v}{(c_2 \fatsemi^\ell c_4)})}
  }{}
  \funrule{applyMiddle(v,l,m)}{
    \rOOfail{l}
  }{
    \sidecond{\neg v \smile m}
  }
  \end{array}
  \]

>>>>

  Value typing \fbox{$v : T$}
  \begin{gather*}
  \dots \qquad
  \inference{
    v : I
  }{
    \vOOcast{v}{\cOOcast{I}{l}{\TOOdyn}} : \TOOdyn
  }
  \quad
  \inference{
    v : P_1 &
    P_1 \smile P_2
  }{
    \vOOcast{v}{\cOOcast{P_1}{l}{P_2}} : P_2
  }
  \end{gather*}
        \caption{Definition of machine state and auxiliary data
          structures.}
        \label{fig:state}
\end{figure}

@@

  \fbox{$applyCast(v,c) = r$}
  \[
  \begin{array}{rclr}
  \funrule{
    applyCast(v,\cOOcast{\star}{l}{\star})
  }{
    \rOOsucc{v}
  }{}
  \funrule{
    applyCast(\vOOcast{v}{\cOOcast{P_1}{l_1}{\star}},\cOOcast{\star}{l_2}{P_2})
  }{
    applyCast(v,\cOOcast{P_1}{l_2}{P_2})
  }{}
  \funrule{
    applyCast(v,\cOOcast{P}{l}{\star})
  }{
    \rOOsucc{\vOOcast{v}{\cOOcast{P}{l}{\star}}}
  }{}
  \funrule{
    applyCast(v,\cOOcast{P_1}{l}{P_2})
  }{
    \rOOsucc{\vOOcast{v}{\cOOcast{P_1}{l}{P_2}}}
  }{\sidecond{P_1 \smile P_2}}
  \funrule{
    applyCast(v,\cOOcast{P_1}{l}{P_2})
  }{
    \rOOfail{l}
  }{\sidecond{P_1 \not\smile P_2}}

@@

  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{?^l}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1}{\bot^l}
  }{
    \sidecond{
       m_1 \not\smile m_2
      \; \text{and} \;
      \forall l. t \neq \bot^{l}
    }
  }

\comprulel{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{\epsilon}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1}{\bot^l}
}{
\sidecond{
  m_1 \not\smile m_2
  \; \text{and} \;
  \forall l. t \neq \bot^{l}
}
}
\comprule{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{?^l}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1 \fatsemi^{l} m_2}{t_2}
}{
\sidecond{
  m_1 \smile m_2
  \; \text{and} \;
  \forall l. t \neq \bot^{l}
}
}
  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{\epsilon}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1 \fatsemi^{\ell} m_2}{t_2}
  }{
    \sidecond{
      m_1 \smile m_2
      \; \text{and} \;
      \forall l. t \neq \bot^{l}
    }
  }
  \end{array}
  \]

@@

\figref{fig:HC-D} defines the functions $id$, $seq$, and $cast$ that
construct \lazyD{} hypercoercions, towards satisifying the Cast ADT.
\figref{fig:HC-D} also defines some auxiliary functions.

The $seq$ operator is defined in terms of coercion composition, which
is the key to compressing coercions and maintaining space efficiency.
A composition operator for coercions typically requires the target
type of the first coercion to match the source type of the second.
However, it is useful to relax this restriction for the D blame
tracking strategy.
%
We shall need optional blame labels, written $\ell$ , that range over
$\epsilon$ or $l$. Then we write composition of hypercoercions as $c_1
\fatsemi^\ell c_2$.  When both $c_1$ and $c_2$ are $\hyperCoercionI$,
their composition is also $\hyperCoercionI$.  If the head of $c_2$ is
a projection, the result is $c_2$.  Otherwise, we need a label to
build the projection. Since the head of $c_2$ is $\epsilon$, its
source type must be a pretype. Thus we know $\ell$ must be a label and
we put it in the projection. When $c_1$ ends with a failure, the
composition is $c_1$ itself. In all the remaining cases, $c_1$ does
not end with a failure.  When $m_1$ and $m_2$ have different top
constructors ($m_1 \not\smile m_2$), the result is a failure coercion,
which needs a label. When $c_2$ starts with a projection, it is to
blame for casting a value to a shallowly inconsistent type. When $c_2$
starts with a no-op, $\ell$ must be a label. And we blame it for
composing shallowly inconsistent hypercoercions. The last two cases
compose $m_1$ and $m_2$ with the auxilliary function $m \fatsemi^\ell
m$, which assumes its inputs have the same top constructor. The
definition of $m \fatsemi^\ell m$ is a straightforward structural
recursion. Going back to the last two cases of $c_1 \fatsemi^\ell
c_2$. When $c_2$ starts with a projection, we do not know whether the
target type of $m_1$ matches the source of $m_2$, although we know
they are shallowly consistent. So we use the $l$ in the projection to
compose the middles $m_1 \fatsemi^l m_2$. When $c_2$ starts with a
no-op, we compose the middles using $\ell$.

The definitions of $id(T)$ and $cast(T_1,l,T_2)$ are straightforward,
although $cast(T_1,l,T_2)$ is unusual in its use of $id$ and
composition.

The proof of correctness for hypercoercions (in
Section~\ref{secc:framework:monoid-correct}) relies on the following
two basic properties of hypercoercions.

\begin{proposition}[\lazyD\ hypercoercions form a monoid]
  \label{thm:hc-monoid}
  For all $c : T_1 \Longrightarrow T_2$,
  $c_1 : T_1 \Longrightarrow T_2$,
  $c_2 : T_2 \Longrightarrow T_3$, and
  $c_3 : T_3 \Longrightarrow T_4$,
  \begin{enumerate}
    \item $seq(id(T_1),c) = c$,
    \item $seq(c,id(T_2)) = c$, and
    \item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  Part (1) and (2) are by induction on $c$. Part (3) is by induction
  on $c_2$ and case analysis on $c_1$ then $c_3$.
\end{proof}

\begin{proposition}[Identity Casts]
  For all $T$ and $l$, $cast(T,l,T) = id(T) $
\end{proposition}
\begin{proof}
  By induction on T.
\end{proof}

@@

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)$] constructs an identity cast from a type
  \item[$seq(c_1,c_2)$] composes two casts
  \item[$cast(T_1,l,T_2)$] constructs a cast from $T_1$ to $T_2$
  \item[$applyCast(v,c)$] applies a cast to a value, producing a result
  \end{description}
  We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

@@

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item if $v : T$, then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item if $\judgeType{v}{T_1}$,
    $ \judgeTypeFT{c_1}{T_1}{T_2}$, and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} >>= f & = & f(v) \\
      \rOOfail{l} >>= f & = & \rOOfail{l}
    \end{array}
    \]
  \item if $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item if $v : \star$, 
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item if $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item if $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item if $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
  \item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

\funrule{
  appS(
    \hcvOOcons{v_1}{c_1}{v_2}{c_2},
    \lsOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}
}{
  \rOOsucc{(\hcvOOcons{v_1}{c_1'}{v_2}{c_2'})}
}{
\\ &&
\text{where} \;
c_1' = seq(c_1,cast(T_1,l,T_3))
\\ &&
\text{and} \;
c_2' = seq(c_2,cast(T_2,l,T_4))
}
\funrule{
  appS(\hcvOOinl{v}{c},\lsOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}})
}{
  \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})}
}{}
\funrule{
  appS(\hcvOOinr{v}{c},\lsOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}})
}{
  \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})}
}{}
\funrule{
  appS(v,\cOOcast{P_1}{l}{P_2})
}{
  \rOOfail{l}
}{\sidecond{P_1 \not\smile P_2}}

@@

  \fbox{$applyMiddle(v,\ell,m) = r$}
  \[
  \begin{array}{rclr}
  \funrule{applyMiddle(\hcvOOtt,\ell,\POOunit)}{\rOOsucc{\hcvOOtt}}{}
  \funrule{applyMiddle(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}\ell,\POOfun{c_3}{c_4})}{
    \rOOsucc{(\hcvOOfun{(c_3 \fatsemi^\ell c_1)}{\rho}{x}{e}{(c_2 
    \fatsemi^\ell c_4)})}
  }{}
  \funrule{applyMiddle(\hcvOOcons{v_1}{c_1}{v_2}{c_2},\ell,\POOprod{c_3}{c_4})}{
    \rOOsucc{(\hcvOOcons{v_1}{(c_1 \fatsemi^\ell c_3)}{v_2}{(c_2 
    \fatsemi^\ell c_4)})}
  }{}
  \funrule{applyMiddle(\hcvOOinl{v}{c_1},\ell,\POOsum{c_3}{c_4})}{
    \rOOsucc{(\hcvOOinl{v}{(c_1 \fatsemi^\ell c_3)})}
  }{}
  \funrule{applyMiddle(\hcvOOinr{v}{c_2},\ell,\POOsum{c_3}{c_4})}{
    \rOOsucc{(\hcvOOinr{v}{(c_2 \fatsemi^\ell c_4)})}
  }{}
  \funrule{applyMiddle(v,l,m)}{
    \rOOfail{l}
  }{
    \sidecond{v \not\smile m}
  }
  \end{array}
  \]
